---
title: "Flipbooks are a blast!"
author: "Evangeline Reynolds"
date: "7/22/2020"
output: html_document
---


```{r, echo = F}
knitr::opts_chunk$set(echo = F)
```


Have you heard of code-movies or "flipbooks"? Maybe not? I'm happy to tell you about this mode of communication and a tool that can help you build them {flipbookr}.  The idea of flipbooks is to show you details of how to get from point A to point B using a data manipulation, analysis, or visualization pipeline.   Often when using tools like R markdown or Jupytr notebooks, we usually only see the input and output for a pipeline of steps.   Having the inputs and outputs nearby is really helpful in communicating the big picture of what is being accomplished with a chunk of code.   

But newcomers and even seasoned practitioners might have trouble figuring out what individual steps in a pipeline accomplish.   This is where flipbooks come in! They seek to illuminate *what's going on in the in-between steps.*   Flipbooks do this by showing in-between outputs as the code is built up.   

The 'flipbook' technique is one that you see from time to time in code presentations.  Presenters realize that showing a code pipeline as it is built will help audiences.

Yet putting together this code-movie experience can be time consuming for creators. There's a lot of copy-paste that has to happen to create the right partial code sequence -- and you can mix yourself up in trying to coordinate it (I've been there!)

The package [{flipbookr}](https://github.com/EvaMaeRey/flipbookr)'s goal is to assist in creating these easy-to-follow-along-with step-by-step experiences of flipbooks.   

So what is {flipbookr} doing? First, it parses entire code chunks as inputs. Then, it identifies good break points for showing output. Given the pause points, it then constructs partial builds of code. Finally then these partial builds are delivered alongside their associated outputs to a slide show tool ([xaringan](https://github.com/yihui/xaringan)!). This gives you the movie-like experience. You can more directly relate a specific function to the change in output it results in.   There is so much cool decision making packed in code pipelines --- the {flipbookr} project is about appreciating those decisions and bringing them into the spotlight.   

---

The nuts and bolts of getting started with flipbookr can be found in this [Easy Flipbook Recipes](https://evamaerey.github.io/flipbooks/flipbook_recipes#1) guide. But it might also be useful to hear about the experience of designing flipbooks from relative newcomers perspective.  Four of my students were kind enough to contribute their flipbooks and also offer some comments on the experience of writing them.  

<!-- So writing code for to create flipbooks is a little different than writing code only to accomplish a task.  It is more like choreographing an experience.  As much as possible, the creators should try to write the code so that their is feedback as code is revealed.    -->




> Conner's experience:

```{r}
xaringanExtra::embed_xaringan("AUS_InteractMap/AUS_tmap.html")
```



> Britt's experience:

```{r}
xaringanExtra::embed_xaringan("magick/magick.html")
```

> Luca's experience:


```{r}
xaringanExtra::embed_xaringan("unemployment/unemployment.html")
```



```{r}
xaringanExtra::embed_xaringan("Idaho Mapping/Idaho_mapping.html")
```


> *The process of producing a flipbook pushed me to think differently about both data wrangling and data visualization. It required me to be more deliberate in how I wrote and ordered my code... - Rachel*

<!-- It required me to be more deliberate in how I wrote and ordered my code, and it revealed redundancies and other inefficiencies in my script. The process also deepened my understanding of the commands that I employed by allowing me to see the output of each individual line of code. -->