---
title: 'Building code-movies with {flipbookr}'
author:
  - Gina Reynolds
  - Rachel Goodman
  - Luca Picci
  - Conner Surrency
  - Brit Woodrum
date: '2021-01-01'
categories:
  - learn
description: |
  here
slug: flipbooks
photo:
  author: Alex Litvin
  url: https://unsplash.com/photos/MAYsdoYpGuk
---


```{r, echo = F}
knitr::opts_chunk$set(echo = F, message = F, warning = F)
```


Have you heard of "code-movies" or "code-flipbooks"? Maybe not?  This blog post will tell you what they are, point you to tools to build them, and showcase student work and their insights about building flipbooks. 

I'll use the terms 'code-movies' and 'flipbooks' interchangeably. 

# What are flipbooks?

Flipbooks help you communicate how to get from 'A' to 'B' using data manipulation, analysis, or visualization code pipelines.   When using R markdown or Jupytr notebooks, we often see the input and output for a pipeline of steps.   Having the inputs and outputs nearby is really helpful in communicating the big picture of what is being accomplished with a chunk of code.  







But newcomers and even seasoned practitioners might have trouble figuring out what the individual steps in a pipeline accomplish.   This is where flipbooks come in! They seek to illuminate what’s going on in *each*  step of the pipeline or plot. Flipbooks do this by showing the *within*-pipeline output states as the code is built up.  


Here's an example. Mouse over to see the movie effect; click to open new tab with original flipbook which lets you move at your own pace and copy code.  

```{r an_mtcars_flipbook, echo = F, eval = F}
set.seed(12345)
library(tidyverse)
mtcars %>% 
  rownames_to_column(var = "model") %>% 
  ggplot() +
  aes(x = wt * 1000) +
  aes(y = factor(cyl)) +
  ggthemes::theme_solarized() +
  geom_boxplot(fill = "snow2") +
  geom_jitter(height = .2, alpha = .8,
              color = "plum3", size = 3) +
  labs(y = "Number of Cylinders") + 
  labs(x = "Weight in pounds") +
  labs(title = "How do number of cylinders and model weight relate?") +
  labs(subtitle = "Motor Trend US magazine's comparison of thirty-two 1973-74 models") +
  theme(plot.title.position = "plot") +
  geom_label(data = tibble(x = 1), # one row dataframe
             fill = "grey85", x = 4600, 
             y = 1.2, # cylinder is a factor w 3 levels
             label = "These observations aren't part of the 'whiskers' of the Tukey boxplot. Outliers are observations further than 1.5 IQR from the nearer of the upper and lower hinge" %>% 
               str_wrap(25)) +
  geom_curve(data = tibble(x = 1),
             curvature = -.2, 
             arrow = arrow(length = 
                             unit(0.1, 
                                  "inches")),
             x = 4550, xend = 5150,
             y = 1.85, yend = 2.8,
             )
```



```{r, eval = F, echo = F}
url <- "../an_mtcars_flipbook/an_mtcars_flipbook_embed.html"
xaringanBuilder::build_gif(url, fps = 2, output_file = "assets/an_mtcars_flipbook_embed.gif")
xaringanBuilder::build_png(url, slides = 17, output_file = "assets/an_mtcars_flipbook_embed.png")

url <- "../Idaho_Mapping/Idaho_mapping.html"
xaringanBuilder::build_gif(url, fps = 2, output_file = "assets/Idaho_mapping.gif")
xaringanBuilder::build_png(url, slides = 43, output_file = "assets/Idaho_mapping.png")

url <- "../AUS_InteractMap/AUS_tmap.html"
xaringanBuilder::build_gif(url, fps = 2, output_file = "assets/AUS_tmap.gif")
xaringanBuilder::build_png(url, slides = 5, output_file = "assets/AUS_tmap.png")

url <- "../magick/magick.html"
xaringanBuilder::build_gif(url,  fps = 2, output_file = "assets/magick.gif")
xaringanBuilder::build_png(url, slides = 18, output_file = "assets/magick.png")

url <- "../unemployment/unemployment.html"
xaringanBuilder::build_gif(url,  fps = 2, output_file = "assets/unemployment.gif")
xaringanBuilder::build_png(url, slides = 42, output_file = "assets/unemployment.png")

library(magrittr)
list.files(path = "assets", pattern = "\\.pdf$") %>% 
  paste0("assets/", .) ->
pdfs_in_assets

unlink(pdfs_in_assets)
```



```{r}
create_html_use_static_and_gif <- function(static_file,
                                           gif_file,
                                           href = "https://github.com/EvaMaeRey/flipbookr",
                                           width = 475,
                                           height = width*9/16){

  paste0(
    '<a href="',
    href,
    '" target="_blank">',
    '<img class="static" width = ', width, ', height = ', height,  ', src="',
    static_file,
    '"><img class="active" width = ', width, ', height = ', height,  ', src="',
    gif_file,
    '">',
    '</a>')

}


create_html_use_static_and_gif("assets/an_mtcars_flipbook_embed.png", 
                               "assets/an_mtcars_flipbook_embed.gif",
                               href = "https://evamaerey.github.io/rstudio_education_blog/an_mtcars_flipbook/an_mtcars_flipbook_embed.html", width = 700) -> 
example_reveal

create_html_use_static_and_gif("assets/unemployment.png", 
                               "assets/unemployment.gif", 
                               href = "https://evamaerey.github.io/rstudio_education_blog/unemployment/unemployment.html"
                               ) -> 
unemployment_reveal

create_html_use_static_and_gif("assets/magick.png", 
                               "assets/magick.gif", 
                               href = "https://evamaerey.github.io/rstudio_education_blog/magick/magick.html") -> 
magick_reveal

create_html_use_static_and_gif("assets/Idaho_mapping.png", 
                               "assets/Idaho_mapping.gif",
                               href = "https://evamaerey.github.io/rstudio_education_blog/Idaho_Mapping/Idaho_mapping.html") -> 
idaho_reveal

create_html_use_static_and_gif("assets/AUS_tmap.png", 
                               "assets/AUS_tmap.gif",
                               href = "https://evamaerey.github.io/rstudio_education_blog/AUS_InteractMap/AUS_tmap.html") ->  
australia_reveal

#### write css static and hover ####

write_css_static_hover <- function(){

"```{css}
.static {
  background: white;
}
.static:hover {
  opacity:0;
}
```"

}

my_css <- write_css_static_hover()

# try <- viztoc:::flipbook_webshots_build_and_use_gif(dir = "magick", base_url = "magick/magick.html", num_pages = 10)
# 
# webshot::webshot(url = "magick/magick.html")

```

```{css}
.static {
  position:absolute;
  background: white;
  }
.static:hover {
  opacity:0;
  }
```

`r example_reveal`

<!-- <div style="position:relative;padding-top:56.25%;"> -->
<!-- <iframe src="https://evamaerey.github.io/rstudio_education_blog/an_mtcars_flipbook/an_mtcars_flipbook_embed.html" frameborder="2" webkitallowfullscreen mozallowfullscreen allowfullscreen style="position:absolute;top:0;left:0;width:100%;height:100%;" allowtransparency="true"> -->
<!-- </iframe> -->
<!-- </div> -->


---

<br><br>

### Code-movies in recent presentations

The 'flipbook' or code-movie technique is one that you may have seen in coding presentations.  Presenters know that showing a code pipeline as it is built helps audiences digest workflows.  So it's worth choreographing a set of slides to break down the build. If the alignment between slides is good --- and if we don't have disruptive slide transitions (wipes, spins, fades) --- we get to enjoy a little movie: the coordinated evolution of code and output.  Some examples are here:

- [*Forecasting* - Mitchelle O'Hara Wild](https://youtu.be/DhDOTxojQ3k?t=350){target="_blank"}
- [*A Gentle Guide to the Grammar of Graphics with ggplot2* - Garrick Aden-Buie](https://pkg.garrickadenbuie.com/gentle-ggplot2/#42){target="_blank"}
- [*3D mapping, plotting, and printing with rayshader* - Tyler Morgan-Wall](https://youtu.be/sB8CYGlPN0o?t=158){target="_blank"}


### Building flipbooks fast and reliably with {flipbookr}

While code-movies deliver helpful insight to consumers, putting together this code-movie experience can be time consuming. There's a lot of copy-paste that has to happen to create the right partial code sequence -- and you can mix yourself up trying to coordinate it (I've been there!).

The [{flipbookr}](https://github.com/EvaMaeRey/flipbookr){target="_blank"} package's goal is to help create these easy-to-follow-along-with, step-by-step experiences --- without the copy and paste pain!  So all you need to do is write your pipeline once -- then you can let {flipbookr} take over and create the coordinated evolution of code and output movie to communicate the process.

Basically {flipbookr}, with the rmarkdown slide show package [xaringan](https://github.com/yihui/xaringan){target="_blank"}, does three things: 

1. Parses an Rmd code pipeline from the chunk you indicate (you name the chunk)
2. Identifies good break points in that code chunk pipeline; the default is finding balanced parentheses at the ends of lines.
3. Spawns a bunch of code chunks with these *partial builds of code*, separated by slide breaks.
4. Displays partial code and builds side by side and sequentially on html slides, giving us a movie-like experience . 

There is so much decision-making packed into our code pipelines.  The {flipbookr} project is about making it easy to bringing those decisions to light so they can be appreciated, examined, discussed, relived!


### Taking {flipbookr} for a spin

After installing flipbookr `install.packages("flipbookr")` there are a couple of ways you might get started.

One way is using the [Easy Flipbook Recipes](https://evamaerey.github.io/flipbooks/flipbook_recipes#1){target="_blank"} guide.  You'll put together a basic flipbook with step-by-step instructions. 

<!-- there's also a [video companion guide](https://www.youtube.com/watch?v=07xEB4q8bXo&feature=youtu.be) to the recipes book.  -->


Alternatively, you can use the "A minimal flipbook" template that comes with the package.  Using R Studio, with the installed package you can request the basic template with `File -> New File -> Rmarkdown -> A Minimal Flipbook`.

<!-- In the future, we hope to deliver a new mode of creating embedded mini flipbooks from *within Rmarkdown files* themselves. You will be able to create instant mini flipbooks with the function `flipbookr::embed_flipbook()`. You (or your students!) will be able to create flipbooks with less fuss than ever before!  Stay tuned or, if you are feeling brave and helpful, check out the [functions](https://github.com/EvaMaeRey/flipbookr/blob/master/R/h_write_instant_flipbook.R) or [examples template](https://github.com/EvaMaeRey/flipbookr/blob/master/inst/rmarkdown/templates/flipbook-embed/skeleton/skeleton.Rmd).   -->



### What to expect: insights and products using {flipbookr} 

<!-- {.tabset .tabset-fade .tabset-pills} -->

Before you start building your own flipbooks, it might also be useful to see some examples from some other folks.  

Four of my graduate research assistants at the [Unversity of Denver's Korbel School of International Studies](https://korbel.du.edu/){target="_blank"}, with the support of an RConsortium Grant, have built several excellent flipbooks here, showcasing {gganimate}, {ggplot2} mapping, {tmap} and {magick}.  





<!-- So writing code for to create flipbooks is a little different than writing code only to accomplish a task.  It is more like choreographing an experience.  As much as possible, the creators should try to write the code so that their is feedback as code is revealed.    -->

Rachel demos how to [build maps with {ggplot2}](https://evamaerey.github.io/rstudio_education_blog/Idaho_Mapping/Idaho_mapping.html){target="_blank"} looking at the political landscape in Idaho.  She works a bunch with thematic elements with the theme() function, displaying the effect of each thematic decision step-by-step. 

`r idaho_reveal`

<!-- <div style="position:relative;padding-top:56.25%;"> -->
<!-- <iframe src="https://evamaerey.github.io/rstudio_education_blog/Idaho_Mapping/Idaho_mapping.html" frameborder="2" webkitallowfullscreen mozallowfullscreen allowfullscreen style="position:absolute;top:0;left:0;width:100%;height:100%;" allowtransparency="true"> -->
<!-- </iframe> -->
<!-- </div> -->


Rachel also contributed this comment about how building plots with the intent of showing in a flipbook differed from the usual build:


> *The process of producing a flipbook pushed me to think differently about both data wrangling and data visualization. It required me to be more deliberate in how I wrote and ordered my code, and it revealed redundancies and other inefficiencies in my script. The process also deepened my understanding of the commands that I employed by allowing me to see the output of each individual line of code.*

Conner [explores the {tmap} package](https://evamaerey.github.io/rstudio_education_blog/AUS_InteractMap/AUS_tmap.html){target="_blank"} showing city population sizes in Australia.  He also does a cool dive into world map projections, cycling through a bunch of projection options.  

`r australia_reveal` 


<!-- <div style="position:relative;padding-top:56.25%;"> -->
<!-- <iframe src="https://evamaerey.github.io/rstudio_education_blog/AUS_InteractMap/AUS_tmap.html" frameborder="2" webkitallowfullscreen mozallowfullscreen allowfullscreen style="position:absolute;top:0;left:0;width:100%;height:100%;" allowtransparency="true"> -->
<!-- </iframe> -->
<!-- </div> -->

Brit demos [the {magick} package](https://evamaerey.github.io/rstudio_education_blog/magick/magick.html){target="_blank"} showing unfolding image manipulation pipelines and combining images.

`r magick_reveal`


<!-- <div style="position:relative;padding-top:56.25%;"> -->
<!-- <iframe src="https://evamaerey.github.io/rstudio_education_blog/magick/magick.html" frameborder="2" webkitallowfullscreen mozallowfullscreen allowfullscreen style="position:absolute;top:0;left:0;width:100%;height:100%;" allowtransparency="true"> -->
<!-- </iframe> -->
<!-- </div> -->

Luca [visualizes changes in youth unemployment in Europe](https://evamaerey.github.io/rstudio_education_blog/unemployment/unemployment.html){target="_blank"} 
first by faceting in {ggplot2} by year, and then using {gganimate}. 

<!-- <div style="position:relative;padding-top:56.25%;"> -->
<!-- <iframe src="https://evamaerey.github.io/rstudio_education_blog/unemployment/unemployment.html" frameborder="2" webkitallowfullscreen mozallowfullscreen allowfullscreen style="position:absolute;top:0;left:0;width:100%;height:100%;" allowtransparency="true"> -->
<!-- </iframe> -->
<!-- </div> -->


`r unemployment_reveal` 


---

### Acknowledgements

The {flipbookr} project has been fortunate to build on incredible tools, in particular {xaringan} and the Rmarkdown. It's inspired by the amazing tools {flipbookr} aims to demo --- particularly ggplot2 and tidyverse data manipulation pipelines. 

Lots of folks have helped build {flipbookr}, especially Emi Tanaka and Garrick Aden-Buie, with visionary insights and ground-breaking work on the code-movie experience and automated code-movie builds in {xaringan}.  

---

### Projects to watch

Related to the flipbooks project is {codehover} and shiny-flipbooks!  






